---
title: "Using the `powers` package"
author: "Vincenzo Coia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the `powers` package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

__Note__: This package isn't actually meant to be "serious". I made it just for teaching purposes.

Want to compute powers without having to use `^` or `**`? Give `sqrt` some friends by loading the `powers` package! 

```{r}
library(powers)
```

## Basic square, cube or reciprocal transformations

Want squares, cubes, or reciprocals? Easy.

```{r}
square(1:10)
cube(1:10)
reciprocal(1:10)
```

`apply`-family friendly, too!

```{r}
my_list <- list(1:10, 0.5, -0.7)
## So base-R-boring!
lapply(my_list, function(x) x^2)
## Use powers instead!
lapply(my_list, square)
```

It even comes with the ability to display its output in plot form!

```{r}
reciprocal(1:10, plot_it=TRUE)
```

Wow!!!

## Beyond the basic transformations - introducing boxcox.transform

The boxcox transformation is based on the old `box.cox()` function from the car package ([see here for that function's documentation](https://www.rdocumentation.org/packages/car/versions/1.1-1/topics/box.cox)). Basically, what want to do is to compute a box-cox transformation of a variable, which is a type of power transformation. The box-cox transformation depends on the lambda (λ) parameter value(s). The box-cox function is defined as x' = (x^λ - 1) / λ when λ ≠ 0 and x' = log x when λ = 0. 

Let's see the box-cox function in action!

```{r}
boxcox.transform(x = 1:10, lambda = 4)
boxcox.transform(x = 2:8, lambda = -20)
boxcox.transform(x = 8:16, lambda = 0)
```

Additionally, lambda may be a vector! If lambda is a vector of values, then the output will be a matrix of values where each column corresponds to a different lambda value (and is named to indicate what lambda is being used).

```{r}
boxcox.transform(x = 1:10, lambda = c(0,2))
```

That example was quite simple. Let's try out bunch of different negative positive values in a lambda vector.

You can see with the below example that the function can be quite powerful. You only need to edit one line to produce box-cox transformations of a data vector that use different lambda values. Furthermore, you are able to view and further manipulate all the transformed data values that correspond to the different lambda values in one handy matrix. 

```{r}
boxcox.transform(x = 1:10, lambda = c(-100,-10,-5,-2,-1,0,1,2,5,10,100))
```

Note that the values of the vector, x, that we want to transform under boxcox.transform must be positive. I have made it so that if x is a vector of negative values the minimum constant will be added so that the x vector values are all positive. 

```{r}
boxcox.transform(x = -5:2, lambda = 4)
```

So, you can see why I added a third parameter called makepos to the boxcox.transform function. This makepos parameter lets you input a value to make the vector that you input positve. 

```{r}
boxcox.transform(x = -5:2, lambda = 4, makepos = 9)
```

If you set makepos to a value such that the x vector of values is still negative, I have made the function so that it will automatically add the minimum constant to every x value to make all the x values positive. Also, I made it so that a warning message indicating as such will be displayed for the person who is trying to use the function.

Let's try out that rookie mistake.

```{r}
boxcox.transform(x = -5:2, lambda = 4, makepos = 2)
```

Clearly, in the above function, when makepos is 2, adding 2 to -5,-4,-3,or -2 isn't enough to make those values positive. So, that's where the minimum automatic addition of the smallest constant to make all the x values positive comes into play. 

By what I mentioned above, we would expect the above function would yield the same output as when we don't have a makepos argument at all.

```{r}
boxcox.transform(x = -5:2, lambda = 4)
```

As we expected, when makepos is an insufficient number to make all the x values positive, the output of the boxcox.transform() function is the same when you have that makepos argument and when you don't.
